\ifdefined\MAINTEXT \else Supplemental Material: \\ \fi Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2017-10-26 almo-langevin.tex
Changed 3N -> 3M in SI equation 6
Major modifications of the main text to take into account recent literature

% define variable that compiles the main text (as opposed to SI text)
%\newcommand*{\MAINTEXT}{}

@@

\title{
\ifdefined\MAINTEXT
\else
Supplemental Material: \\
\fi
Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
%%for weakly-coupled atoms
%% for non-covalently bonded atoms
}

@@

\begin{abstract} 
A robust method for constructing compact localized orbitals in density functional theory enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems. High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces. An application to liquid water demonstrates that extremely low computational overhead of this procedure makes it suitable for routine medium-size simulations whereas its linear scaling cost allows to extend accurate first-principle molecular dynamics to previously inaccessible time and length scales. 
\end{abstract}

@@

Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
\emph{ab initio} molecular dynamics (AIMD) has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids. %[of gas-phase molecules and condensed-phase systems.] 
Unfortunately, the computational cost of the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms, which severely limits the \emph{length scales} accessible by AIMD. %and hinders it application to large heterogeneous systems (interfaces, nuclei, defects). 
To address this issue, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.

% RZK: search and include recent references
In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors
~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}. 
Most LS methods~\cite{a:ls-rev-1999,a:ls-rev-2012} explore the natural locality of the one-electron density matrix (DM). 
However, in typical DFT calculations that require diffuse basis functions, the DM approach is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis. 
Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.

% DM-based LS methods: a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign

To obviate the DM optimization, another set of electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the development of orbital-based LS methods has been all but abandoned because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.

@@

Thus, the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets. 
On the length and time scales accessible in typical AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.

In this work, we developed a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center and a predefined localization radius $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is a set of all local Gaussian atomic orbitals of one molecule. 
However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, RZK:10.1016/j.jcp.2011.11.032}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second step, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first step. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}. 
Since the importance of electron transfer decays exponentially with distance, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems. %with just a few local variables. 
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method for accurate AIMD simulations of large molecular systems.

@@

The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals. First, while ALMOs are variationally optimized in both SCF steps, the occupied subspace defined in the first step must remain fixed during the second step to ensure convergence. Second, neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$. In addition to these issues characteristic to ALMOs, the variational optimization in any AIMD method is never truly complete and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$. While these nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations, they accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure. 
Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$. 
%This however would dramatically increase the computational cost of the ALMO AIMD. %as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver.

In this work, we propose another approach that obviates the need in a coupled-perturbed solver, relaxes tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus enables us to maintain stable dynamics and to keep the complexity and cost of simulations low. 
%
In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense nonvariational component of the forces. 
The difference between these approximate ALMO forces and the \emph{reference} forces that could be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
%
\begin{align}
\label{eq:deltaf}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$. $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization. 
$\delta f_{i\alpha} (t)$ can be reduced to zero \emph{systematically} by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$.

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{RZK:PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}. 
Inspired by this method, ALMO AIMD is chosen to be governed by the Langevin equation of motion that is first written in terms of the unknown reference forces:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise: %with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
% 
The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at specified temperature $T$~\cite{a:Kubo-1986}. 
%will generate the correct Maxwell-Boltzmann distribution of velocities for a desired temperature $T$~\cite{a:Kubo-1986}. 
Larger $\gamma$ lead to faster equilibration but more distorted results for dynamical properties. %In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled.

%It has previously been shown that Langevin dynamics can be utilized to correct inaccurate force calculations to maintain correct dynamical properties.~\cite{a:langevin-why,a:2ndcpmd,b:tuckerman-stat,a:ceriotti}

%By utilizing Eq.\ (\ref{eq:langevin}) in conjunction with the ALMO SCF method of force calculation, we will show that discontinuity in the PES can be ignored for sufficiently large $\gamma$. 
% Perhaps this statement is not correct: Note that increasing $\gamma$ increases the sample size required to generate accurate dynamical properties, as more random Gaussian noise is introduced.

%The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
%%During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
%Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{\text{SCF}}$. 
%Previously, $\epsilon_{\text{SCF}}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
%By increasing the allowable $\epsilon_{\text{SCF}}$, a great reduction in computational cost can be achieved. 
%Obviously, this introduces a significant error into the calculation of forces during each time step.

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD. %and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.

@@

%RZK: There is a problem with the integral equation below. The r.h.s. depends on $i$ but in the l.h.s. does not. If I remember correctly we evaluate the integral for molecules, not atoms. In liquid water all our molecules are the same and we do not have to deal with the issue of different Delta for different molecules. Eliminating the problem
In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) weighed over all types of atoms with different $m_i$
%
\begin{align}
\label{eq:delta}
\Delta &= (6 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$ .
%%
%\begin{align}
%\label{eq:eqipartition}
%\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
%\end{align}

>>>>

% define variable that compiles the main text (as opposed to SI text)
\newcommand*{\MAINTEXT}{}

@@

\title{
\ifdefined\MAINTEXT
\else
Supplemental Material: \\
\fi
Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
%%for weakly-coupled atoms
%% for non-covalently bonded atoms
%RZK: without one-electron density matrix
}

@@

\begin{abstract} 
We coupled a robust method for constructing compact localized orbitals within the framework of Kohn-Sham density functional theory with a tunable Langevin equation of motion to obviate the commonly employed optimization of the one-electron density matrix and to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems. 
%An application to liquid water demonstrates that the linear scaling complexity and low computational overhead of the developed method make it suitable for accurate first-principle simulation of a wide range of systems: from routine medium-size systems to previously inaccessible length scales.
%well suited for routine first-principle modeling as well as 
%for a wide range of length scales -- from routinely accessible to previously inaccessible.
%In contrast to linear-scaling first-principle simulation methods that exploit the locality of the one-electron density matrix we coupled a robust method for constructing compact localized orbitals with a modified Langevin equation of motion to create an efficient \emph{ab initio} molecular dynamics method for weakly-interacting systems.
%A robust method for constructing compact localized orbitals in density functional theory is coupled with a tunable Langevin equation of motion to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems. 
An application to liquid water demonstrates that the low computational overhead of the developed method makes it well-suited for routine simulations whereas its linear scaling complexity allows to extend first-principle dynamical studies of molecular systems to previously inaccessible length scales. 
%
%%%A robust method for constructing compact localized orbitals in density functional theory enables us to create low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems. 
%%%High efficiency of simulations is achieved by calculating the forces approximately without lengthy optimization of localized orbitals and by adjusting the Langevin equation of motion to stabilize the dynamics with the imperfect forces. 
%%%An application to liquid water demonstrates that a low computational overhead of this procedure makes it suitable for routine medium-size simulations whereas its linear scaling cost allows to extend accurate first-principle molecular dynamics 
%%%%[RZK: of molecular systems] 
%%%to previously inaccessible length scales. 
%%%% RZK: no DM, only orbitals. Make it sound like a theory development not implementation focused.
\end{abstract}

@@

Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
\emph{ab initio} molecular dynamics (AIMD) has become an important tool to study processes in molecules and materials.
%weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids. 
Unfortunately, the computational cost of the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms, which severely limits the \emph{length scales} accessible by AIMD. %and hinders it application to large heterogeneous systems (interfaces, nuclei, defects). 
To address this issue, substantial efforts have been directed to the development of linear-scaling (LS) DFT.

In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors. 
%~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}. 
Most LS methods~\cite{a:ls-rev-1999, a:ls-rev-2012, Kussmann2013, Aarons2016, optimal-basis-too} explore the natural locality of the one-electron density matrix (DM). 
However, with accurate multifunction basis sets the DM DFT becomes advantageous only for impractically large systems~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:ls-almo}.
%Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014, more}. 
This issue is rectified in optimal-basis DM methods~\cite{Skylaris2005, Nakata2015, Mohr2015} that contract large basis sets into a small number of new localized functions and then optimize the DM in the contracted basis. 
%Later: Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} 
%While most large-scale DFT calculations are performed using this class of methods
Despite becoming the most popular approach to LS DFT the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{convergence-gilan}. 
%
%Richters2014 -- DOI:10.1063/1.4869865
%Aarons2016 -- DOI:10.1063/1.4972007
%Kussmann2013 -- DOI:10.1002/wcms.1138
%Vitale2017 -- DOI:10.1063/1.4978684
%Skylaris2005 -- DOI:10.1063/1.1839852
%Hine2011 -- DOI:10.1103/PhysRevB.83.195102
%Bowler2010 -- DOI:10.1088/0953-8984/22/7/074207
%Nakata2015 -- DOI:10.1039/C5CP00934K
%Krajewski -- PRB, 73, 041105R
%Kolafa2003 -- J Comput Chem 25: 335–342, 2004
%
From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved without costly DM optimization and with significantly fewer variables. %especially in the case of large accurate basis sets. 
Advantages of the orbitals-only DFT are especially pronounced in accurate calculations that require many basis functions per atom.
%From this point of view, it is desirable to obviate the DM optimization and use another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
%
% direct optimization of such compact localized nonorthogonal orbitals
%
%To obviate the DM optimization, another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
%From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:review1999, a:weitao-yang-2013,japanese-guys} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013,japanese-guys-citedinmy2013paper}.

@@

Thus, despite impressive progress of LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign} the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, RZK:10.7567/JJAP.55.1102B1,Hine2011, Bowler2010}. 
%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign} ... 
%Thus, the high computational overhead of existing LS DFT methods restrict their use in [dynamical] AIMD simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, RZK:10.7567/JJAP.55.1102B1,Hine2011, Bowler2010}. 
On typical length and time scales required in practical and accurate AIMD simulations, LS DFT still cannot compete with the straightforward low-cost cubically-scaling KS DFT.

In this work, we present an AIMD method that overcomes difficulties of orbital-only local DFT to achieve LS with extremely low computational overhead. 
%Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
To demonstrate advantages of the new method we applied it here to systems of weakly-interacting molecules. 
However, the same approach is readily applicable to systems of strongly-interacting fragments that do not form strong covalent bonds such as ionic salts, liquids, and semiconductors. 
A generalization of the method to all finite-gap systems, including covalently bonded atoms, will be reported later. 
%It is important to emphasize that delocalized electrons in metals, which are particularly challenging for all LS DFT~\cite{Kuehne,recent-Skylaris}, cannot be described efficiently within the presented framework.

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, RZK:10.1016/j.jcp.2011.11.032}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}. 
Since the importance of electron transfer decays exponentially with distance in finite-gap materials~\cite{a:ls-rev-1999}, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems. %with just a few local variables. 
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method for accurate AIMD simulations of large molecular systems.

@@

The challenge of adopting ALMO DFT for dynamical simulations arises from the slightly nonvariational character of the localized orbitals. First, while ALMOs are variationally optimized in both SCF stages, the occupied subspace defined in the first stage must remain fixed during the second stage to ensure convergence. Second, neglected electron transfer effects can suddenly become active in the course of a dynamical simulation when a neighboring molecule passes the localization threshold $R_{c}$. In addition to these issues specific to ALMO DFT, the variational optimization in any AIMD method is never truly complete in practice
%[exact convergence is never achieved in practical calculation] 
and interrupted once the maximum norm of the gradient of the energy with respect to the electronic descriptors drops below small but nevertheless finite convergence threshold $\epsilon_{\text{SCF}}$. These nonvariational errors do not affect the accuracy of static ALMO DFT calculations, geometry optimization, and Monte-Carlo simulations. Unfortunately they tend to accumulate in molecular dynamics trajectories leading to non-physical sampling and eventual failure. 
Traditional strategies to cope with these problems are computationally expensive and include computing the nonvariational contribution to the forces via a variational coupled-perturbed procedure~\cite{Kussmann2013}, increasing $R_c$, and decreasing $\epsilon_{\text{SCF}}$. 
%This however would dramatically increase the computational cost of the ALMO AIMD. %as longer-distance but physically unimportant interactions must be calculated with an additional iterative coupled perturbed solver.

In this work, we propose another approach that obviates the need in a coupled-perturbed solver, relaxes tight constraints on $R_{c}$ and $\epsilon_{\text{SCF}}$, and thus enables us to maintain stable dynamics and to keep the algorithmic complexity and cost of simulations low. 
%
In our approach, the forces on atoms are calculated \emph{approximately} after the two-stage ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense nonvariational component of the forces. 
The difference between these approximate ALMO forces and the \emph{reference} forces that could be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
%
\begin{align}
\label{eq:deltaf}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$. $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization. 
$\delta f_{i\alpha} (t)$ can be reduced to zero \emph{systematically} by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$.

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{RZK:10.1002/wcms.1176}. 
Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise: %with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
% 
The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at a specified temperature $T$~\cite{a:Kubo-1986}. 
In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled. 
%Increasing $\gamma$ leads to faster equilibration but more distorted results for dynamical properties.

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD.

@@

%RZK: There is a problem with the integral equation below. The r.h.s. depends on $i$ but in the l.h.s. does not. If I remember correctly we evaluate the integral for molecules, not atoms. In liquid water all our molecules are the same and we do not have to deal with the issue of different Delta for different molecules. Eliminating the problem
In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) averaged over atoms with different $m_i$
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \frac{1}{3} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
In practice, we found (see results below) that this approach is not particularly accurate because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.
%%
%\begin{align}
%\label{eq:eqipartition}
%\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
%\end{align}

Thus, the inherently stochastic approach presented here does not aim to produce fully time-reversible dynamics for atomic nuclei. 
However, with $\gamma=0$, the nuclei will follow time-reversible Newtonian trajectories sufficiently closely as long as partially optimized ALMOs remain close to the ground state and keep $\Delta$ small. 
%
%(for systems with slight discontinuities in the potential energy surfaces)
%
Although not a required component of ALMO AIMD, the predictor part of the Kolafa scheme~\cite{kolafa2003} adopted to localized orbitals was found to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps.

%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.
